### 2019-03-28 paper/paper.tex
Further work on paper

\documentclass[9pt]{entcs}

\usepackage{cleveref}
\usepackage{entcsmacro}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{bussproofs}
\usepackage{stmaryrd}
\usepackage{cleveref}

\sloppy
% The following is enclosed to allow easy detection of differences in
% ascii coding.
% Upper-case    A B C D E F G H I J K L M N O P Q R S T U V W X Y Z
% Lower-case    a b c d e f g h i j k l m n o p q r s t u v w x y z
% Digits        0 1 2 3 4 5 6 7 8 9
% Exclamation   !           Double quote "          Hash (number) #
% Dollar        $           Percent      %          Ampersand     &
% Acute accent  '           Left paren   (          Right paren   )
% Asterisk      *           Plus         +          Comma         ,
% Minus         -           Point        .          Solidus       /
% Colon         :           Semicolon    ;          Less than     <
% Equals        =3D           Greater than >          Question mark ?
% At            @           Left bracket [          Backslash     \
% Right bracket ]           Circumflex   ^          Underscore    _
% Grave accent  `           Left brace   {          Vertical bar  |
% Right brace   }           Tilde        ~

\newenvironment{bprooftree}
{\leavevmode\hbox\bgroup}
{\DisplayProof\egroup}

\crefname{lemma}{lemma}{lemmata}
\Crefname{lemma}{lemma}{lemmata}

% Setting modes
\newcommand{\type}[1]{\textsc{#1}}
\newcommand{\term}[1]{\mathsf{#1}}
\newcommand{\constructor}[1]{\mathbf{#1}}
\newcommand{\category}[1]{\textsc{#1}}
\newcommand{\functortxt}[1]{\mathsf{#1}}
\newcommand{\nattranstxt}[1]{\mathsf{#1}}
\newcommand{\function}[1]{\mathsf{#1}}
\newcommand{\morphism}[1]{\mathsf{#1}}

% HoTT
\newcommand{\deprod}[3]{\prod_{#1 : #2} #3} % dependent product
\newcommand{\hset}{\type{hSet}} % sets
\newcommand{\0}{\textbf{0}} % empty type
\newcommand{\1}{\textbf{1}} % unit type
\newcommand{\inl}{\term{inl}} %left inclusion
\newcommand{\inr}{\term{inr}} % right inclusion
\newcommand{\prl}{\term{pr}_1} % first projection
\newcommand{\prr}{\term{pr}_2} % second projection
\newcommand{\pairTm}[2]{(#1 , #2)} % pairing
\newcommand{\weq}{\simeq} % weak equivalence
\newcommand{\setquot}[2]{#1 / #2} % quotient type
\newcommand{\transport}[2]{\function{transport}^{#1} #2} % transport
\newcommand{\classof}{\constructor{class}} % class construcor
\newcommand{\classeq}{\constructor{classeq}} % path between classes
\newcommand{\isaprop}[1]{\function{isaprop}(#1)} % a type is a set
\newcommand{\isaset}[1]{\function{isaset}(#1)} % a type is a set
\newcommand{\trunct}[1]{\lVert #1 \rVert} % propositional truncation
\newcommand{\truncel}[1]{| #1 |} % constructor of truncation
\newcommand{\depeq}[3]{#2 =_{#1} #3} % dependent equality type

% Category theory
\newcommand{\idm}[1]{\morphism{id}_{#1}} % identity morphism
\newcommand{\compm}[2]{#1 \circ #2} % composition of morphisms
\newcommand{\iso}[2]{#1 \cong #2} % isomorphisms
\newcommand{\setoids}{\category{Setoid}} % category of setoids
\newcommand{\functor}[2]{#1 \longrightarrow #2} % functors
\newcommand{\idf}[1]{\functortxt{id}_{#1}} % identity functor
\newcommand{\Cf}[1]{\functortxt{C}_{#1}} % constant functor
\newcommand{\compf}[2]{#1 \circ #2} % composition of functors
\newcommand{\prodf}[2]{#1 \times #2} % product of functors
\newcommand{\sumf}[2]{#1 + #2} % sum of functors
\newcommand{\nattrans}[2]{#1 \Longrightarrow #2} % natural transfomations
\newcommand{\ctrans}[1]{\nattranstxt{c}_{#1}} % constant transformation
\newcommand{\constrt}[1]{\nattranstxt{constr}_{#1}} % constructor transformation
\newcommand{\idt}[1]{\nattranstxt{id}} % identity transformation
\newcommand{\compt}[2]{#1 \circ #2} % composition of natural transformations
\newcommand{\lwhisker}[2]{#1 \vartriangleleft #2} % left whiskering of transformations
\newcommand{\inlt}{\nattranstxt{inl}} %left inclusion of transformations
\newcommand{\inrt}{\nattranstxt{inr}} % right inclusion of transformations
\newcommand{\prlt}{\nattranstxt{pr}_1} % first projection of transformations
\newcommand{\prrt}{\nattranstxt{pr}_2} % second projection of transformations
\newcommand{\pairt}[2]{(#1 , #2)} % pairing of transformations
\newcommand{\idtoiso}{\function{idtoiso}} % identity to iso

% Polynomials
\newcommand{\poly}{\mathcal{P}} % type of polynomials
\newcommand{\C}{\constructor{C}} % constant polynomial
\newcommand{\I}{\constructor{I}} % identity polynomial
\newcommand{\sumP}[2]{#1 + #2} % sum of polynomials
\newcommand{\prodP}[2]{#1 \times #2} % product of polynomials

% Equivalence relations
\newcommand{\pathR}[1]{\function{Path} \> #1} % path equivalence relation
\newcommand{\sumR}[2]{#1 + #2} % sum of equivalence relations
\newcommand{\prodR}[2]{#1 \times #2} % product of equivalence relations
\newcommand{\polyR}[2]{#1 \> #2} % action of polynomials on equivalence relations

% Endpoints
\newcommand{\ep}[3]{\mathcal{E}_{#1}(#2,#3)} % type of endpoinst
\newcommand{\id}[1]{\constructor{id}_{#1}} % identity endpoints
\newcommand{\comp}[2]{#1 \cdot #2} % composition of endpoints
\newcommand{\inle}{\constructor{inl}} %left inclusion of endpoints
\newcommand{\inre}{\constructor{inr}} % right inclusion of endpoints
\newcommand{\prle}{\constructor{pr}_1} % first projection of endpoints
\newcommand{\prre}{\constructor{pr}_2} % second projection of endpoints
\newcommand{\pair}[2]{(#1 , #2)} % pairing of endpoints
\newcommand{\Ce}{\constructor{c}} % constant endpoint
\newcommand{\constr}{\constructor{constr}} % constructor endpoint

% Signatures
\newcommand{\pt}[0]{\textsf{pt}}
\newcommand{\pthI}[0]{\textsf{pth}}
\newcommand{\pthA}[0]{\textsf{arg}}
\newcommand{\pthlh}[0]{\textsf{lhs}}
\newcommand{\pthrh}[0]{\textsf{rhs}}
\newcommand{\pointc}[1]{#1_{\pt}} % point constructor
\newcommand{\pathI}[1]{#1_{\pthI}} % path constructor index
\newcommand{\pathA}[1]{#1_{\pthA}} % path constructor index
\newcommand{\pathlh}[1]{#1_{\pthlh}} % left endpoint of path constructor
\newcommand{\pathrh}[1]{#1_{\pthrh}} % left endpoint of path constructor

\newcommand{\comm}{\function{comm}}
\newcommand{\trunc}{\function{trunc}}

% Algebras
\newcommand{\semP}[1]{\llbracket #1 \rrbracket} % semantics of polynomials
\newcommand{\prealg}[1]{\category{PreAlg}_{\hset}(#1)} % category of prealgebras
\newcommand{\forget}[1]{F_{#1}} % forgetful functor
\newcommand{\semE}[1]{\llbracket #1 \rrbracket} % semantics of endpoints

\newcommand{\Alg}[1]{\category{Alg}_{\hset}(#1)} % category of algebras
\newcommand{\constrA}[1]{\function{map}_{#1}} % point constructor of algebra
\newcommand{\constrP}[1]{\function{path}_{#1}} % path constructor of algebra

% Algebras in setoids
\newcommand{\semPT}[1]{\langle #1 \rangle} % semantics of polynomials in setoids
\newcommand{\semET}[1]{\langle #1 \rangle} % semantics of endpoints in setoids
\newcommand{\prealgst}[1]{\category{PreAlg}_{\setoids}(#1)} % category of setoid prealgebras
\newcommand{\Algst}[1]{\category{Alg}_{\setoids}(#1)} % category of algebras in setoids

% Displayed algebras
\newcommand{\polydact}[2]{\overline{#1} \> #2}
\newcommand{\epdact}[2]{\overline{#1} \> #2}
\newcommand{\polydmap}[2]{\overline{#1} \> #2}

\def\lastname{Please list Your Lastname Here}
\begin{document}
\begin{frontmatter}
  \title{An Example Paper} \author{My
    Name\thanksref{ALL}\thanksref{myemail}}
  \address{My Department\\ My University\\
    My City, My Country} \author{My Co-author\thanksref{coemail}}
  \address{My Co-author's Department\\My Co-author's University\\
    My Co-author's City, My Co-author's Country} \thanks[ALL]{Thanks
    to everyone who should be thanked} \thanks[myemail]{Email:
    \href{mailto:myuserid@mydept.myinst.myedu} {\texttt{\normalshape
        myuserid@mydept.myinst.myedu}}} \thanks[coemail]{Email:
    \href{mailto:couserid@codept.coinst.coedu} {\texttt{\normalshape
        couserid@codept.coinst.coedu}}}
\begin{abstract} 
  This is a short example to show the basics of using the ENTCS style
  macro files.  Ample examples of how files should look may be found
  among the published volumes of the series at the ENTCS Home Page
  \texttt{http://www.elsevier.com/locate/entcs}.
\end{abstract}
\begin{keyword}
  Please list keywords from your paper here, separated by commas.
\end{keyword}
\end{frontmatter}
\section{Introduction}
\label{intro}

\section{Preliminaries}
\label{sec:prelim}
We start by recalling some definitions from homotopy type theory (HoTT) and fixing the notation.
The first definition we need, is the \emph{dependent equality type}.

\begin{definition}
\label{def:depeq}
Given a type $X$, a type family $Y$ on $X$, and a path $p : x = y$, then we have a map
\[
\transport{Y}{p} : Y \> x \rightarrow Y \> y.
\]
Given a path $p : x = y$ and inhabitants $z_1 : Y \> x$ and $z_2 : Y \> y$, we write $\depeq{p}{z_1}{z_2}$ for $\transport{Y}{p} \> z_1 = z_2$.
\end{definition}

One of the core features of HoTT, is that equality is proof relevant.
This means that not every inhabitant of $x = y$ is necessarily equal.
There might be some types for which this is the case though, and we call such types \emph{sets}.
More precisely, we define

\begin{definition}
\label{def:prop}
\label{def:set}
A type $X$ is a \emph{(mere) proposition} if for all $x, y : X$ we have $x = y$.
A type $X$ is a \emph{set} if for all $x, y : X$ the type $x = y$ is a proposition.
We write $\isaprop{X}$ and $\isaset{X}$ to say that $X$ is a proposition and set respectively.
\end{definition}

To construct higher inductive types, we make use of only two.
These are the \emph{quotient type} and the \emph{propositional truncation}.
For their elimination and computation rules, we refer the reader to the literature.

\begin{definition}
\label{def:quot}
Let $X$ be a type and let $R$ be an equivalence relation on $X$.
The \emph{quotient type} $\setquot{X}{R}$ is the higher inductive type generated by the following rules
\begin{center}
\begin{bprooftree}
\AxiomC{$x : X$}
\UnaryInfC{$\classof \> x : \setquot{X}{R}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$x, y : X$}
\AxiomC{$r : R \> x \> y$}
\BinaryInfC{$\classeq \> r : \classof \> x = \classof \> y$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$\isaset{\setquot{X}{R}}$}
\end{bprooftree}
\end{center}
\end{definition}

\begin{definition}
\label{def:trunc}
Let $X$ be a type.
The \emph{propositional truncation} $\trunct{X}$ is the higher inductive type generated by the following rules
\begin{center}
\begin{bprooftree}
\AxiomC{$x : X$}
\UnaryInfC{$\truncel{x} : \trunct{X}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$\isaprop{\trunct{X}}$}
\end{bprooftree}
\end{center}
\end{definition}

Next we briefly discuss category theory in univalent foundations.
We start by recalling the definition of categories.
This is almost the same as the usual definition in mathematics, but their is a slight discrepancy between the objects and arrows.
While the objects can be any type, the arrows have to form a set.
If equality was proof relevant, then this would give the structure of a higher category instead of an ordinary one.
This leads to the following definition.

\begin{definition}
\label{def:cat}
A \emph{category} $\mathcal{C}$ consists of
\begin{itemize}
	\item a type $\mathcal{C}_0$ of objects;
	\item for each $X, Y : \mathcal{C}_0$ a set $\mathcal{C}_1(X,Y)$ of morphisms;
	\item for each $X : \mathcal{C}_0$ a morphisms $\idm{X}$;
	\item for all objects $X, Y, Z : \mathcal{C}_0$ and morphisms $f : \mathcal{C}_1(X,Y)$ and $g : \mathcal{C}_1(Y,Z)$, a morphism $\compm{f}{g} : \mathcal{C}_1(X,Z)$
\end{itemize}
such that the usual associativity and identity laws holds.
\end{definition}

We define isomorphisms in the usual way and we denote the type of isomorphisms from $X$ to $Y$ by $\iso{X}{Y}$.
... is the notion of \emph{univalent categories}.
These are categories in which equality on objects is equivalent to isomorphisms between them.
More precisely, we define them as follows.

\begin{definition}
\label{def:univalent}
Let $\mathcal{C}$ be a category.
Note that for all objects $X$ and $Y$ we have a map $\idtoiso_{X,Y} : X = Y \rightarrow \iso{X}{Y}$ sending the reflexivity path to the identity isomorphism.
Then we say $\mathcal{C}$ is \emph{univalent} if $\idtoiso_{X,Y}$ is a weak equivalence for each $X$ and $Y$.
\end{definition}

The primary example of a univalent category is the category $\hset$ of sets and functions between them.
Its univalence follows from the univalence axiom.
Another example of a univalent category which we use frequently, is the category of \emph{setoids}.

\begin{definition}
\label{def:setoid}
We define
\begin{itemize}
	\item A \emph{setoid} is a set $X$ together with an equivalence relation on $X$;
	\item A \emph{setoid morphism} between two setoids  is a map between the underlying sets which preserves the equivalence relation.
\end{itemize}
The \emph{category of setoids} $\setoids$ is the category with setoids and setoid morphisms as objects and morphisms.
\end{definition}

If $X$ is a set and $R$ is an equivalence relation on $X$, then we write $(X, R)$ for the setoid with this data.
When it is clear from the context, we write $\equiv$ for the relation.
We finish this section by giving some operations of functor and natural transformations.
For categories $\mathcal{C}$ and $\mathcal{D}$, we write $\functor{\mathcal{C}}{\mathcal{D}}$ for the type of functors from $\mathcal{C}$ to $\mathcal{D}$.
For functor $F, G : \functor{\mathcal{C}}{\mathcal{D}}$, we write $\nattrans{F}{G}$ for the type of natural transformations from $\mathcal{F}$ to $\mathcal{G}$.

\begin{lemma}
\label{def:functor}
We have the following functors.
\begin{itemize}
	\item For each category $\mathcal{C}$, we have the identity $\idf{\mathcal{C}} : \functor{\mathcal{C}}{\mathcal{C}}$.
	\item For $F : \functor{\mathcal{C}_1}{\mathcal{C}_2}$ and $G : \functor{\mathcal{C}_2}{\mathcal{C}_3}$, we have a composition $\compf{F}{G} : \functor{\mathcal{C}_1}{\mathcal{C}_2}$;
	\item If $\mathcal{D}$ has binary products, then for $F, G : \functor{\mathcal{C}}{\mathcal{D}}$, we have a product $\prodf{F}{G} : \functor{\mathcal{C}}{\mathcal{D}}$;
	\item If $\mathcal{D}$ has binary sum, then for $F, G : \functor{\mathcal{C}}{\mathcal{D}}$, we have a sum $\sumf{F}{G} : \functor{\mathcal{C}}{\mathcal{D}}$;
	\item For each object $X : \mathcal{D}$, we have a constant functor $\Cf{X} : \functor{\mathcal{C}}{\mathcal{D}}$.
\end{itemize}
\end{lemma}

\begin{lemma}
\label{def:nattrans}
We have the following transformations.
\begin{itemize}
	\item Suppose, $F, G : \functor{\mathcal{C}}{\mathcal{D}}$
	\[
	\inlt : \nattrans{F}{\sumf{F}{G}}
	\quad \quad
	\inrt : \nattrans{G}{\sumf{F}{G}}
	\quad \quad
	\prlt : \nattrans{\prodf{F}{G}}{F}
	\quad \quad
	\prrt : \nattrans{\prodf{F}{G}}{G}
	\]
	\item Given two transformations $\eta_1 : \nattrans{\compf{F}{G_1}}{\compf{F}{G_2}}$ and $\eta_2 : \nattrans{\compf{F}{G_1}}{\compf{F}{G_3}}$, we have a pairing 
	\[
	\pairt{\eta_1}{\eta_2} : \nattrans{\compf{F}{G_1}}{\compf{F}{\prodf{G_2}{G_3}}}.
	\]
	\item Given functors $F : \functor{\mathcal{C}_1}{\mathcal{C}_2}$ and $G_1. G_2 : \functor{\mathcal{C}_2}{\mathcal{C}_3}$, and a transformation $\eta : \nattrans{G_1}{G_2}$, we have
	\[
	\lwhisker{F}{\eta} : \nattrans{\compf{F}{G_1}}{\compf{F}{G_2}}
	\]
	\item Given a functor $F : \functor{\mathcal{C}}{\hset}$, a set $X : \hset$, and $x : X$, we have a transformation $\ctrans{x} : \nattrans{F}{\Cf{X}}$
\end{itemize}
\end{lemma}

\section{Signature of HITs}
\label{sec:signature}
Before we study the construction of set truncated HITs, we first give a precise definition of those.
We do this by internalizing a scheme, so, more concretely, we define a type of signatures.
In this section, we give the definition of signatures, and then in the upcoming sections, we define the notion of HITs for a signature.

Higher inductive types are freely generated by both point and path constructors.
The data of the point constructor is a \emph{polynomial functor}.
These are described by the following type.

\begin{definition}
\label{def:poly}
We define the type $\mathcal{\poly}$ of \emph{polynomial functors} inductively by the following rules
\begin{center}
\begin{bprooftree}
\AxiomC{$X : \hset$}
\UnaryInfC{$\C \> X : \poly$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$\I : \poly$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P : \poly$}
\AxiomC{$Q : \poly$}
\BinaryInfC{$\sumP{P}{Q} : \poly$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P : \poly$}
\AxiomC{$Q : \poly$}
\BinaryInfC{$\prodP{P}{Q} : \poly$}
\end{bprooftree}
\end{center}
\end{definition}

Next we describe the possible path constructors.
Note that these depend on the point constructor.
Furthermore, the path constructors are universally quantified equations of which the arguments depend polynomially on the type being defined.
Again we use an inductive definition to describe all possibilities.

\begin{definition}
\label{def:endpoint}
Given polynomials, $A, P, Q : \poly$, we define the type $\ep{A}{P}{Q}$ of \emph{endpoints} inductively by the following rules
\begin{center}
\begin{bprooftree}
\AxiomC{$P : \poly$}
\UnaryInfC{$\id{A} : \ep{A}{P}{P}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P, Q, R : \poly$}
\AxiomC{$e_1 : \ep{A}{P}{Q}$}
\AxiomC{$e_2 : \ep{A}{Q}{R}$}
\TrinaryInfC{$\comp{e_1}{e_2} : \ep{A}{P}{R}$}
\end{bprooftree}
\end{center}
\begin{center}
\begin{bprooftree}
\AxiomC{$P, Q : \poly$}
\UnaryInfC{$\inle : \ep{A}{P}{\sumP{P}{Q}}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P, Q : \poly$}
\UnaryInfC{$\inre : \ep{A}{Q}{\sumP{P}{Q}}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P, Q : \poly$}
\UnaryInfC{$\prle : \ep{A}{\prodP{P}{Q}}{P}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P, Q : \poly$}
\UnaryInfC{$\prre : \ep{A}{\prodP{P}{Q}}{Q}$}
\end{bprooftree}
\end{center}
\begin{center}
\begin{bprooftree}
\AxiomC{$\constr : \ep{A}{A}{\I}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P : \poly$}
\AxiomC{$X : \hset$}
\AxiomC{$x : X$}
\TrinaryInfC{$\Ce \> x : \ep{A}{P}{\C \> X}$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$P, Q, R: \poly$}
\AxiomC{$e_1 : \ep{A}{P}{Q}$}
\AxiomC{$e_2 : \ep{A}{P}{R}$}
\TrinaryInfC{$\pair{e_1}{e_2} : \ep{A}{P}{\prodP{Q}{R}}$}
\end{bprooftree}
\end{center}
\end{definition}

Next we put it all together.
Note that we index the path constructors by a type meaning that we could possibly have infinitely many path constructors.

\begin{definition}
\label{def:signature}
A \emph{HIT signature} $\Sigma$ consists of
\begin{itemize}
	\item A polynomial $\pointc{\Sigma} : \poly$;
	\item A type $\pathI{\Sigma}$;
	\item A family $\pathA{\Sigma} : \pathI{\Sigma} \rightarrow \poly$;
	\item For each $j : \pathI{\Sigma}$, endpoints $\pathlh{\Sigma} \> j, \pathrh{\Sigma} \> j : \ep{P}{Q \> j}{I}$.
\end{itemize}
\end{definition}

To illustrate the possibilities of this definition, we consider two examples.
The first one is a set with a commutative operation and the second one is the propositional truncation.

\begin{example}
\label{ex:commsig}
Define a signature $\comm$ such that
\begin{itemize}
	\item $\pointc{\comm} = \prodP{\I}{\I}$;
	\item $\pathI{\comm} = \1$;
	\item $\pathA{\comm} = \prodP{\I}{\I}$;
	\item $\pathlh{\comm} \> j = \comp{\constr}{\id{\prodP{\I}{\I}}}$;
	\item $\pathrh{\comm} \> j = \comp{\constr}{\pair{\comp{\prle}{\id{\prodP{\I}{\I}}}}{\comp{\prre}{\id{\prodP{\I}{\I}}}}}$;
\end{itemize}
\end{example}

Similarly, one can define signatures for groups or finite sets.
Since signatures can depend on types, we can define the propositional truncation.

\begin{example}
\label{ex:truncsig}
Let $A$ be a set.
Define a signature $\trunc$ such that
\begin{itemize}
	\item $\pointc{\trunc} = \C \> A$;
	\item $\pathI{\trunc} = \1$;
	\item $\pathA{\trunc} = \prodP{\I}{\I}$;
	\item $\pathlh{\trunc} \> j = \comp{\prle}{\id{\prodP{\I}{\I}}}$;
	\item $\pathrh{\trunc} \> j = \comp{\prre}{\id{\prodP{\I}{\I}}}$;
\end{itemize}
\end{example}

\section{Algebras}
\label{sec:algebras}
Now we know what signatures are, the next goal is to define HITs for signatures.
For that we need to give its introduction, elimination, and computation rules.
In this section, we study the introduction rule of HITs and we do that by defining a category of algebras on a polynomial.
Furthermore, since ultimately we want to construct HITs as quotients, we do not just define algebra in sets, but also in setoids.
Let us start by interpreting polynomials as functors.

\begin{definition}
\label{def:sempoly}
For each $P : \poly$, we define a functor $\semP{P} : \functor{\hset}{\hset}$ as follows
\begin{itemize}
	\item $\semP{\C \> X} = \Cf{X}$;
	\item $\semP{\I} = \idf{\hset}$;
	\item $\semP{\sumP{P}{Q}} = \sumf{\semP{P}}{\semP{Q}}$;
	\item $\semP{\sumP{P}{Q}} = \prodf{\semP{P}}{\semP{Q}}$.
\end{itemize}
\end{definition}

Now we define the category $\prealg{\Sigma}$ to be the category of algebras on the functor $\semP{\pointc{\Sigma}}$.
Its objects are maps $\semP{\pointc{\Sigma}} \> X \rightarrow X$.
We call this the category of \emph{prealgebras} since these do not yet satisfy the equations in $\Sigma$.
Note that we always have a forgetful functor $\forget{P} : \functor{\semP{P}}{\hset}$.

To obtain actual algebras of $\Sigma$, we need to be able to interpret the equations and for that, we first need to interpret endpoints.
These are interpreted as natural transformations, which are constructed using the transformations from Lemma \ref{def:nattrans}.
Beside those, we need a transformation $\constrt{P} : \nattrans{\comp{\forget{P}}{\semP{P}}}{\comp{\forget{P}}{\semP{\I}}}$ for each polynomial $P$.
Its components are given by the prealgebra map.

\begin{definition}
\label{def:semendpoint}
For each endpoint $e : \ep{A}{P}{Q}$, we define a natural transformation $\semE{e} : \nattrans{\compf{\forget{A}}{\semP{P}}} {\compf{\forget{A}}{\semP{Q}}}$
\begin{itemize}
	\item $\semE{\id{P}} = \idt{\compf{\forget{A}}{\semP{P}}}$;
	\item $\semE{\comp{e_1}{e_2}} = \compt{\semE{e_1}}{\semE{e_2}}$;
	\item $\semE{\inle} = \lwhisker{\forget{A}}{\inlt}$;
	\item $\semE{\inre} = \lwhisker{\forget{A}}{\inrt}$;
	\item $\semE{\prle} = \lwhisker{\forget{A}}{\prlt}$;
	\item $\semE{\prre} = \lwhisker{\forget{A}}{\prrt}$;
	\item $\semE{\pair{e_1}{e_2}} = \pairt{\semE{e_1}}{\semE{e_2}}$;
	\item $\semE{\Ce \> t} = \ctrans{t}$;
	\item $\semE{\constr} = \constrt{A}$.
\end{itemize}
\end{definition}

Now we have everything in place to define algebras on $\Sigma$.
An algebra on $\Sigma$ need to have an operation which satisfies certain equations.
For the operation, we use $\prealg{\pointc{\Sigma}}$.
Since the carrier of each prealgebra is a set, the equations form a proposition.
Hence, we define the category of \emph{$\Sigma$-algebras} as a full subcategory of $\prealg{\pointc{\Sigma}}$.

\begin{definition}
\label{def:algebra}
Let $\Sigma$ be a HIT signature.
Then we define the category $\Alg{\Sigma}$ of \emph{$\Sigma$-algebras} as the full subcategory of $\prealg{\pointc{\Sigma}}$ such that each object satisfies
\[
\deprod{j}{\pathI{\Sigma}}{\deprod{x}{\pathA{\Sigma} \> j}{\semE{\pathlh{\Sigma} \> j} \> x = \semE{\pathrh{\Sigma} \> j} \> x}}.
\]
\end{definition}

For an algebra $X$, we denote its operation by $\constrA{X} : \semP{\pointc{\Sigma}} \> X \rightarrow X$.
The path witnessing the equalities of the algebra is denoted by $\constrP{X} : \deprod{j}{\pathI{\Sigma}}{\deprod{x}{\pathA{\Sigma} \> j}{\semE{\pathlh{\Sigma} \> j} \> x = \semE{\pathrh{\Sigma} \> j} \> x}}$.
The category of $\Sigma$-algebras is univalent, which means that isomorphism of algebras corresponds to equality of algebras.
This follows from the fact that the category of algebras on a functor is univalent and that univalence is preserved under taking full subcategories.

\begin{proposition}
\label{prop:univalg}
The category of $\Sigma$-algebras is univalent.
\end{proposition}

Before we look at algebras in setoids in more detail, we recall the examples in the previous section and look what algebras on those signatures are.

\begin{example}
\label{ex:commalg}
Recall the signature $\comm$ from Example \ref{ex:commsig}.
An algebra of $\comm$ consists of a set $X$ and a map $f : X \times X \rightarrow X$ such that for all $(x , y) : X \times X$, we have $f(x, y) = f(y, x)$.
\end{example}

\begin{example}
\label{ex:truncalg}
Recall $\trunc$ from Example \ref{ex:truncsig}.
For a set $A$. an algebra of $\trunc \> A$ consists of a set $X$ and a map $f : A \rightarrow X$ such that for all $(x , y) : X \times X$, we have $x = y$.
In particular, this means $X$ is a proposition.
\end{example}

Recall that the core idea is that we define the initial algebra in sets by taking the quotient of the initial algebra in setoids.
For that reason, we also define algebras in setoids and for these, we take a slightly different approach.
Instead of defining the interpretation of the polynomial $P$ inductively making use of binary sums and products in $\setoids$, we define an equivalence relation on $\semP{P}$.

To define the action of polynomials on equivalence relations, we need so preliminary operations.
For a set $T$, we have an equivalence relation $\pathR{T}$ on $T$ such that $\pathR{T} \> x \> y = (x = y)$.
Furthermore, given types $X$ and $Y$ with equivalence relations $R_X$ and $R_Y$ on them, we can define equivalence relations $\sumR{R_X}{R_Y}$ and $\prodR{R_X}{R_Y}$ on $X + Y$ and $X \times Y$ respectively.
These are defined such that
\[
\prodR{R_X}{R_Y} \> (x_1,y_1) \> (x_2,y_2) = (R_X \> x_1 \> x_2) \times (R_Y \> y_1 \> y_2)
\]
\[
\begin{array}{ccc}
\sumR{R_X}{R_Y} \> (\inl \> x_1) \> (\inl \> x_2) = R_X \> x_1 \> x_2 & \quad & \sumR{R_X}{R_Y} \> (\inr \> y_1) \> (\inr \> y_2) = R_Y \> y_1 \> y_2\\
\sumR{R_X}{R_Y} \> (\inl \> x) \> (\inr \> y) = \0 & \quad & \sumR{R_X}{R_Y} \> (\inr \> y) \> (\inl \> x) = \0\\ 
\end{array}
\]

\begin{definition}
Let $R$ be an equivalence relation on a set $X$ and let $P : \poly$ be a polynomial.
By induction, we define an equivalence relation $\polyR{P}{R}$ on $\semP{P} \> X$.
\begin{itemize}
	\item $\polyR{(\C \> T)}{R} = \pathR{T}$;
	\item $\polyR{\I}{R} = R$;
	\item $\polyR{(\sumP{P}{Q})}{R} = \sumR{(\polyR{P}{R})}{(\polyR{Q}{R})}$;
	\item $\polyR{(\prodP{P}{Q})}{R} = \prodR{(\polyR{P}{R})}{(\polyR{Q}{R})}$.
\end{itemize}
Now we define a functor $\semPT{P} : \functor{\setoids}{\setoids}$ such that $\semPT{P}(X,R) = (\semP{P} \> X, P \> R)$.
\end{definition}

Again this gives rise to a category $\prealgst{\Sigma}$ of \emph{setoid prealgebras on $\Sigma$}.
To define algebras on $\Sigma$, we also need to interpret endpoints.
However, these are just setoid morphisms instead of natural transformations.

\begin{definition}
Let $e : \ep{A}{P}{Q}$ be an endpoint and let $X$ be a setoid prealgebra on $A$.
Then we define $\semE{e}$ to be the setoid morphism from $\semE{P}(X)$ to $\semE{Q}(X)$ whose carrier is $\semP{e}$.
\end{definition}

The category of algebras on $\Sigma$ is defined differently for setoids that for sets.
While for sets we require actual equalities, for setoids the equations need to be witnessed by the equivalence relation instead.
Note that such relations are families of propositions meaning again this gives rise to a proposition and thus we define i as a full subcategory.

\begin{definition}
Let $\Sigma$ be a HIT signature.
Then we define the category of \emph{$\Sigma$-setoid-algebras} as the full subcategory of $\prealgst{\pointc{\Sigma}}$ such that each object satisfies
\[
\deprod{j}{\pathI{\Sigma}}{\deprod{x}{\pathA{\Sigma} \> j}{\semE{\pathlh{\Sigma} \> j} \> x \equiv \semE{\pathrh{\Sigma} \> j} \> x}}.
\]
\end{definition}

Since the category of setoids is univalent, the category of $\Sigma$-setoid-algebras is univalent as well.

\begin{proposition}
The category of $\Sigma$-setoid-algebras is unvialent.
\end{proposition}

\section{The Induction Principle}
\label{sec:induction}
The introduction rules of higher inductive types are covered by the algebra structure, but the elimination and computation rules require additional structure.
The main feature of the elimination rule is that it allows defining dependent ``homomorphisms'' to families with some kind of ``dependent algebra structures'' and the computation rule says the algebraic structure is preserved.
To formulate these two rules, we make use of \emph{displayed algebras}, which represent the input of the elimination rule.
When we got all this machinery in place, we define the notion of \emph{higher inductive type} for a signature.

Since our goal is to construct HITs, we need verify the induction rule and for that, we use initial algebra semantics.
More specifically, we show that the initial algebra satisfies the induction rule.
Hence, to obtain it a HIT, it suffices to construct the initial algebra for which we use category theory.

\subsection{Displayed Algebras}
A displayed algebra is the input of the elimination rule.
This means that we have a dependent family and a dependent map over the point constructor.
Furthermore, dependent versions of the equations in the signature needs to hold.
To formulate these requirements more precisely, we need two preliminary definitions.
The first one gives an action of polynomials on type families while the second one interprets endpoints as dependent maps.

\begin{definition}
Given are a polynomial $P$ and a family $Y : X \rightarrow \hset$.
We define a family $\polydact{P}{Y} : \semP{P} \> X \rightarrow \hset$ by induction
\begin{itemize}
	\item $\polydact{\C \> X}{Y} \> x = X$;
	\item $\polydact{\I}{Y} \> x = Y \> x$;
	\item $\polydact{\sumP{P}{Q}}{Y} \> (\inl \> x) = \polydact{P}{Y} \> x$;
	\item $\polydact{\sumP{P}{Q}}{Y} \> (\inr \> x) = \polydact{Q}{Y} \> x$;
	\item $\polydact{\prodP{P}{Q}}{Y} \> x = \polydact{P}{Y} \> (\prl \> x) \times \polydact{Q}{Y} \> (\prr \> x)$.
\end{itemize}
\end{definition}

\begin{definition}
Let $A$ be a polynomial, $X$ be a prealgebra on $A$, and let $e : \ep{A}{P}{Q}$.
Suppose, we have a family $Y$ on $X$ and a map $c : \deprod{z}{\semP{A}{X}}{\polydact{P}{Y} \> z \rightarrow Y \> (\constrA{X} \> z)}$.
We define a map $\epdact{e}{c} : \deprod{z}{\semP{P}{X}}{\polydact{P}{Y} \> z \rightarrow \polydact{Q}{Y} \> (\semE{e} \> z)}$ by induction on $e$
\begin{itemize}
	\item $\epdact{\id{P}}{c} \> z \> y = y$;
	\item $\epdact{\comp{e_1}{e_2}}{c} \> z \> y = \epdact{e_2}{c} \> (\semE{e} \> X \> z) \> (\epdact{e_1}{c} \> z \> y)$;
	\item $\epdact{\inle}{c} \> z \> y = y$;
	\item $\epdact{\inre}{c} \> z \> y = y$;
	\item $\epdact{\prle}{c} \> z \> y = \prlt \> y$;
	\item $\epdact{\prre}{c} \> z \> y = \prrt \> y$;
	\item $\epdact{\pair{e_1}{e_2}}{c} \> z \> y = (\epdact{e_1}{c} \> z \> y , \epdact{e_2}{c} \> z \> y)$;
	\item $\epdact{\Ce \> t}{c} \> z \> y = t$;
	\item $\epdact{\constr}{c} = c$.
\end{itemize}
\end{definition}

With this in place, we define displayed algebras.
Note that since we are working in a type family, we need to use the dependent equality type from Definition \ref{def:depeq} instead.

\begin{definition}
Let $\Sigma$ be a HIT signature and let $X$ be an algebra on $\sigma$.
Then a \emph{displayed algebra} over $X$ consists of
\begin{itemize}
	\item A type family $Y : X \rightarrow \hset$;
	\item An operation $c_Y : \deprod{z}{\semP{A}{X}}{\polydact{P}{Y} \> z \rightarrow Y \> (\constrA{X} \> z)}$;
	\item For each $j : \pathI{\Sigma}$, $x : \semP{\pathA{\Sigma}} \> X$, and $y : \polydact{\pathA{\Sigma}}{Y} \> x$,  a path $p_Y : \depeq{\constrP{X} \> j \> x}{\epdact{\pathlh{\Sigma}}{c} \> x \> y}{\epdact{\pathrh{\Sigma}}{c} \> x \> y}$.
\end{itemize} 
\end{definition}

Now we got the input for the elimination rule in place and the next step is to look at the output.
This is a dependent map which preserves the algebra structure.
To state this preservation property, we need yet another operation on polynomials.

\begin{definition}
Let $P$ be a polynomial, let $X$ be a set, and let $Y$ be a family of sets on $X$.
Given a map $f : \deprod{x}{X}{Y \> x}$, we define a map $\polydmap{P}{f} : \deprod{x}{\semP{P} \> X}{\polydact{P}{Y} \> x}$ by induction
\begin{itemize}
	\item $\polydmap{\C \> X}{f} \> x = x$;
	\item $\polydmap{\I}{f} = f$;
	\item $\polydmap{\sumP{P}{Q}}{f} \> (\inl \> x) = \polydmap{P}{f} \> x$;
	\item $\polydmap{\sumP{P}{Q}}{f} \> (\inr \> x) = \polydmap{Q}{f} \> x$;
	\item $\polydmap{\prodP{P}{Q}}{f} \> x = (\polydact{P}{f} \> (\prl \> x) , \polydact{Q}{f} \> (\prr \> x))$.
\end{itemize}
\end{definition}

\begin{definition}
Let $Y$ be a displayed algebra over $X$.
Then a displayed algebra map to $Y$ consists of a map $f : \deprod{x}{X}{Y \> x}$ such that for each $x : \semP{\pointc{\Sigma}}{X}$, we have $f(\constrA{X} \> x) = c_Y \> x \> (\polydmap{\pointc{\Sigma}}{f} \> x)$.
\end{definition}

With all this in place, we can define the notion of a \emph{higher inductive type} on a signature.
Note that a HIT needs to be an algebra so that we have the correct introduction rules and note that we use displayed algebras to formulate the elimination principle.

\begin{definition}
Let $\Sigma$ be a HIT signature.
A \emph{higher inductive type} on $\Sigma$ consists of an algebra $H$ such that for each displayed algebra $Y$ on $H$, we have a displayed algebra map to $Y$.
\end{definition}

\subsection{Obtaining Induction from Initiality}
Now suppose that we have some displayed algebra.
Our goal is to construct an algebra whose objects are dependent pairs.

Give the carrier and operation in plain text

\begin{lemma}
pr1 endpoint
\end{lemma}

\begin{lemma}
pr2 endpoint
\end{lemma}

All in all, we get the following construction.

\begin{theorem}
Total algebra and projection
\end{theorem}

\begin{proposition}
Map to total algebra which projects into identity is section.
\end{proposition}

Hence, to construct a HIT, it suffices to construct an initial object in the category of algebras.

\begin{corollary}

\end{corollary}

\section{Constructing the Initial Algebra}

\subsection{Adjunction between Algebras}
To construct the adjunction between the category of algebras, we use two lemmata

\begin{lemma}
Given are categories $\mathcal{C}$ and $\mathcal{D}$, functors $A_{\mathcal{C}} : \functor{\mathcal{C}}{\mathcal{C}}$, $A_{\mathcal{D}} : \functor{\mathcal{D}}{\mathcal{D}}$, and $F : \functor{\mathcal{C}}{\mathcal{D}}$.
Then we get a functor 
\end{lemma}

\begin{lemma}
Given are categories $\mathcal{C}$ and $\mathcal{D}$, functors $A_{\mathcal{C}} : \functor{\mathcal{C}}{\mathcal{C}}$, $A_{\mathcal{D}} : \functor{\mathcal{D}}{\mathcal{D}}$, and an adjunction ....
Suppose, the following diagrams commute

Then we get an adjunction
\end{lemma}

\begin{lemma}
Adjunction between full subcategories
\end{lemma}

Now we apply these to get an adjunction between set HIT-algebras and setoid HIT-algebras.
For this, we do the following steps.

\begin{lemma}
The path setoid commutes with sums and products.

The quotient commutes with sums and products.
\end{lemma}

\begin{lemma}
We have a natural isomorphism (path setoid/quotients commutes with polynomials).
\end{lemma}

Now we can lift the quotient and path setoid.
We can verify the conditions and we get

\begin{lemma}
There is an adjunction between prealgebras ...
\end{lemma}

To show this adjunction lifts to an adjunction between algebras, we need to show that they map algebras to algebras.
We only show this for the path setoid

\begin{lemma}
path setoid endpoint
\end{lemma}

\begin{lemma}
Path setoid factors through algebras
\end{lemma}

All in all, we get

\begin{proposition}
We have an adjunction
\end{proposition}

\subsection{Initial Setoid Algebra}

Hence, we conclude

\begin{theorem}
HITs exist
\end{theorem}

\section{Consequences}

\subsection{Recursion}

\begin{definition}
Let $X$ and $Y$ be algebras on $\Sigma$.
Then we define the \emph{constant displayed algebra} on $X$ from $Y$ as follows
\begin{itemize}
	\item The type family is $Y$ for each $x$;
	\item
\end{itemize}
\end{definition}

\begin{proposition}
Let $X$ be $Y$ be algebras on $\Sigma$.
If we have a displayed algebra from $X$ to the constant displayed algebra on $X$ from $Y$, then we have an algebra homomorphism from $X$ to $Y$.
\end{proposition}

\begin{corollary}
If $X$ is a HIT for $\Sigma$, then for each algebra $Y$ we have an algebra map from $X$ to $Y$.
\end{corollary}

\subsection{Uniqueness of HITs}
A first property of higher inductive types, is that they are unique up to equality.
For that, we first show that higher inductives are initial algebras.
Note that in \Cref{sec:induction} we made use of the converse statement, namely that initial algebras are HITs.

To show a HIT $H$ is initial, we must show that for each algebra $X$ the set of morphisms from $H$ to $X$ is contractible.
This means that there is at most one morphism.
To show the existence, we use the recursion rule.
To show uniqueness, we use an alternative induction principle for families of propositions.
More specifically, we use the following lemma.

\begin{lemma}
Let $H$ be a HIT of $\Sigma$ and suppose $Y$ is a family of propositions on $H$.
If we have an operation
\[
c : \deprod{z}{\semP{A}{X}}{\polydact{P}{Y} \> z \rightarrow Y \> (f \> z)},
\]
then we have a map $\deprod{x}{H}{Y \> x}$.
\end{lemma}

To prove the uniqueness of the map, we use function extensionality and the previous lemma.
This is sufficient to conclude that higher inductive types are initial objects.

\begin{proposition}
If $H$ is a HIT for $\Sigma$, then $H$ is an initial $\Sigma$-algebra.
\end{proposition}

Now we take advantage of the fact that the category of algebras is univalent.
Since initial objects in a univalent category are unique up to equality, we can immediately conclude that HITs are actually unique up to equality.
All in all, we get

\begin{theorem}
If $H_1$ and $H_2$ are HITs for $\Sigma$, the $H_1 = H_2$ as $\Sigma$-algebras.
\end{theorem}

\subsection{Path Spaces of HITs}
The construction of HITs also allows us to characterize the path space up to equivalence.

\begin{proposition}
Let $X$ be a type and let $R$ be an equivalence relation on $X$.
Then for each $x, y : X$, we have an equivalence $R \> x \> y \weq \classof \> x = \classof \> y$.
\end{proposition}

\section{Conclusion}

\begin{thebibliography}{10}\label{bibliography}
\bibitem{cy} Civin, P., and B. Yood, \emph{Involutions on Banach
    algebras}, Pacific J. Math. \textbf{9} (1959), 415--436.

\bibitem{cp} Clifford, A. H., and G. B. Preston, ``The Algebraic
  Theory of Semigroups,'' Math. Surveys \textbf{7}, Amer. Math. Soc.,
  Providence, R.I., 1961.

\bibitem{f} Freyd, Peter, Peter O'Hearn, John Power, Robert Tennent
  and Makoto Takeyama, \emph{Bireflectivity}, Electronic Notes in
  Theoretical Computer Science {\bf 1} (1995), URL:
  \href{https://www.sciencedirect.com/journal/electronic-notes-in-theoretical-computer-science/vol/1/suppl/C}
  {\texttt{http://www.elsevier.com/locate/entcs/volume1.html}}.

\bibitem{em2} Easdown, D., and W. D. Munn, \emph{Trace functions on
    inverse semigroup algebras}, U. of Glasgow, Dept. of Math.,
  preprint 93/52.

\bibitem{r} Roscoe, A. W., ``The Theory and Practice of Concurrency,''
  Prentice Hall Series in Computer Science, Prentice Hall Publishers,
  London, New York (1198), 565pp. With associated web site\\  
  \href{http://www.comlab.ox.ac.uk/oucl/publications/books/concurrency/}
  {\texttt{http://www.comlab.ox.ac.uk/oucl/publications/books/concurrency/}}.

\bibitem{s} Shehadah, A. A., ``Embedding theorems for semigroups with
  involution, `` Ph.D.  thesis, Purdue University, Indiana, 1982.

\bibitem{w} Weyl, H., ``The Classical Groups,'' 2nd Ed., Princeton U.
  Press, Princeton, N.J., 1946.

\end{thebibliography}

\end{document}